---
title: "Data Processing and Exploratory analyses"
author: 
  - "Jake W. Coldiron,"
  - "Zixuan Qiu,"
  - "Tianyuan Deng,"
  - "Aung Nay Win,"
  - "Yuqi Cheng"
output: github_document
always_allow_html: true
---

**Date:** 2023-11-10

```{r message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(plotly)
library(broom)
library(modelr)
library(maps)
```

# Data processing

```{r message=FALSE}
#input the New York HIV data from NYC opendata 
nyc_HIV1=
  GET("https://data.cityofnewyork.us/resource/ykvb-493p.csv",
  query = list("$limit" = 10000)) |>
  content("parsed")|>
  filter(year >= 2017)


nyc_HIV2=
  GET("https://data.cityofnewyork.us/resource/dxnu-p2qd.csv",
  query = list("$limit" = 22000)) |>
  content("parsed")|>
  filter(year >= 2017)
  
```

```{r message=FALSE}
#HIV service resource 
HIV_service=
  GET("https://data.cityofnewyork.us/resource/pwts-g83w.csv",
  query = list("$limit" = 1000)) |>
  content("parsed")|>
  select(unique_id,facilityname,service_type, address,borough,state,zipcode,latitude,longitude)
  
```
**A short  clear description of where and how data were obtained. explain the variables means in the data**
**Calculate the sum, mean, median, standard deviation of HIV/AIDS indicators** 


# Visualizations

**Shows the number of HIV/AIDS diagnoses by race, sex, and age group.**
```{r}
#work
#work
colnames(nyc_HIV1)
colnames(nyc_HIV2)
head(nyc_HIV1)
library(ggplot2)
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

Female <- nyc_HIV1 %>% 
  filter(sex == "Female" & borough == "All" & race !="All")

Male <- nyc_HIV1 %>% 
  filter(sex == "Male"  & borough == "All" & race !="All")

graph1 = bind_rows(Female, Male)

nyc_HIV1$hiv_diagnoses_num <- as.integer(nyc_HIV1$hiv_diagnoses_num)
nyc_HIV2$hiv_diagnoses_num <- as.integer(nyc_HIV2$hiv_diagnoses_num)

ggplot(graph1, aes(x=race, y=hiv_diagnoses_num, fill=sex)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Race", y = "Number of HIV/AIDS Diagnoses", fill = "Sex")

# Graph Age 
age_summary <- nyc_HIV2 %>%
  filter(age != "All" & race == "All" & borough == "All") %>%
  group_by(age) %>%
  summarise(total_hiv_diagnoses = sum(hiv_diagnoses_num))

ggplot(age_summary, aes(x = age, y = total_hiv_diagnoses)) +
  geom_bar(stat = "identity", color = "yellow", width = 0.7) +
  labs(title = "Total HIV Diagnoses by Age Group",
       x = "Age Group",
       y = "Total HIV Diagnoses") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_text(aes(label=total_hiv_diagnoses), vjust=-0.5, color="black", size=3.5)
```
summary:
Both the nyc_HIV1 dataset and the nyc_HIV1 data have 11 variables, nyc_HIV1 is the data related to the number of patients with HIV by gender and nyc_HIV2 is the data related to the number of patients with HIV by age. When analyzing the number of HIV patients by race, the results show that Latino/Hispanic races are used for high prevalence and there are more males than females. When analyzing the number of HIV patients by age group, the results showed that 20-29 year olds had the highest number of total diagnoses and 13-19 year olds had the lowest number of diagnoses.

**Show the map of NYC HIV service resources location.**
```{r}
if (!require(leaflet)) install.packages('leaflet')
library(leaflet)
filtered_data = HIV_service[HIV_service$longitude != 0 & HIV_service$latitude != 0,]
# Create the map
m <- leaflet(filtered_data) %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    popup = ~paste("<strong>Facility Name:</strong>", facilityname,
                   "<br><strong>Service Type:</strong>", service_type,
                   "<br><strong>Address:</strong>", address,
                   "<br><strong>Borough:</strong>", borough,
                   "<br><strong>Zip:</strong>", zipcode)
  )

# Print the map
print(m)
```
summary:
Here, an interactive map was created using the leaflet package to visualize the distribution of HIV service facilities and detailed information to help better understand the allocation of resources for HIV prevention and treatment.

**Show the distribution of the HIV/AIDS diagnoses by neighborhood.** 
```{r}
#work
# I'll see if I can find the log&lat for the neighborhoods
diagnoses <- nyc_HIV1 %>% 
  select(borough, hiv_diagnoses_num)

# Plot bar chart
ggplot(diagnoses, aes(x = reorder(borough, hiv_diagnoses_num), y = hiv_diagnoses_num)) +
  geom_col() +
  labs(x = "Neighborhood",
       y = "Number of HIV Diagnoses",
       title = "HIV Diagnoses by Neighborhood") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```
```{r}
filtered_data = HIV_service[HIV_service$longitude != 0 & HIV_service$latitude != 0,]
# Create the map
m <- leaflet(filtered_data) %>%
  addTiles("https://server.arcgisonline.com/ArcGIS/rest/services/World_Street_Map/MapServer/tile/{z}/{y}/{x}") %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    popup = ~paste("<strong>Facility Name:</strong>", facilityname,
                   "<br><strong>Service Type:</strong>", service_type,
                   "<br><strong>Address:</strong>", address,
                   "<br><strong>Borough:</strong>", borough,
                   "<br><strong>Zip:</strong>", zipcode)
  )
m <- m %>% addCircleMarkers(
  radius = sqrt(nyc_HIV1$hiv_diagnoses_num),
  fillOpacity = 0.4,
  color = "orange",
  stroke = TRUE,
  label = sprintf("HIV Diagnoses: %s", nyc_HIV1$hiv_diagnoses_num),
  labelOptions = labelOptions(
    textOnly = TRUE,
    direction = "auto"
  )
)
m
```
summary:
The number of HIV cases varies by community, with the Staten Island community having a much smaller number of confirmed cases, the Brooklyn community having a larger number of confirmed cases, and the other communities having a similar number of cases.


**Trend analysis:analyze the trend of the number of HIV/AIDS diagnoses over time.**
```{r}
#work
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

# Plot diagnoses over time
ggplot(All, aes(x = year, y = hiv_diagnoses_num)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "HIV Diagnoses Trend (All Sex)",
    x = "Year",
    y = "Number of Diagnoses"
  )
```
summary:
The number of patients diagnosed in different years showed significant differences. The overall trend is decreasing from year to year. This also reflects the growing importance that has been placed on HIV.



# Exploratory statistical analyses.
**Correlation analysis:After the Visualizations is there any association between different variables such as neighborhood, age, gender, and race.**

**T-test for HPV prevalence in gender-specific populations.**
```{r}
female_data <- nyc_HIV1[nyc_HIV1$sex == "Female",]$hiv_diagnoses_num
male_data <- nyc_HIV1[nyc_HIV1$sex == "Male",]$hiv_diagnoses_num
# Perform T test
t.test(female_data, male_data)
```
summary:
The t-test passed the test of significance indicating that there was a significant difference in the number of males and females with HIV.

**Test whether "hiv_diagnoses_num" is significantly different across ethnicities.**
```{r}
anova1 <- anova(lm(hiv_diagnoses_num ~ race, data = nyc_HIV1))
print(anova1)
```
summary:
The ANOVA passed the test of significance, indicating that there is indeed a significant difference in the number of people of different races who have HIV.

**Test whether "hiv_diagnoses_num" is significantly different between age groups.**
```{r}
anova2 <- anova(lm(hiv_diagnoses_num ~ age, data = nyc_HIV2))
print(anova2)
```
summary:
The ANOVA passed the test of significance indicating that there is a significant difference in the number of HIV in different age groups.

**Test whether there is a significant difference between people of different races and different age groups.**
```{r}
race_table <- table(nyc_HIV1$race, nyc_HIV1$sex)
race_table
# Perform chi-square test
chisq.test(race_table)
```
summary:
The chi-square test passes the test of significance, indicating that there is a significant relationship between race and gender.

# linear regression analysis
```{r}
# Defining a Unique Thermal Coding Processor
library(caret)
encoder <- dummyVars(~ borough + sex + race, data = nyc_HIV1)
# Solo thermal coding of data
encoded_data <- predict(encoder, newdata = nyc_HIV1)
# Merging the coded and original data sets
nyc_HIV3 <- cbind(nyc_HIV1, encoded_data)
nyc_HIV4<-nyc_HIV3[,-c(2,3,4,5,7,8,9,10,11,12,18,21)]
model <- lm(hiv_diagnoses_num ~. , data = nyc_HIV4)
summary(model)
```
summary:
The coefficients of all variables passed the significance test, indicating that all of these independent variables have a significant effect on hiv_diagnoses_num. The coefficient before the variable "boroughStaten\\\nIsIand" has the largest absolute value, which means that it has the largest effect on the number of HIV patients. In addition, the coefficients before these variables are all negative, indicating that they all negatively affect hiv_diagnoses_num.Finally, the multiple linear regression model constructed ishiv_diagnoses_num=4514.4981-2.0224*year-354.5227*boroughBronx-358.4969*boroughBrooklyn-362.8923*boroughManhattan-363.2155*boroughQueens-373.2117*boroughStaten\\nIsland-26.8822*sexFemale-6.1453*sexMale-62.6910*raceAsian /Pacific\\nIslander-36.4375*raceBlack-41.7176*raceLatino/Hispanic-65.3393*raceOther/Unknown-57.9514*raceWhite.

# Regression model test
```{r}
#Drawing demonstration
par(mfrow=c(2, 2))  
plot(model)

# Plot of predicted values against residuals
residuals <- resid(model)
fitted_values <- predict(model)
plot(fitted_values, residuals,
     xlab = "Fitted values",
     ylab = "Residuals",
     main = "Residuals vs. Fitted")
abline(h = 0, col = "red")
```

**summary:Here is a test of the effectiveness of the regression model constructed above. Use plot(model) to draw residual plots, scatter plots and other diagnostic plots. Residual plots are used to check for linear properties, heteroskedasticity, outliers, etc. Scatter plots can be used to check for correlation between independent variables, and QQ plots are used to check whether the residuals conform to the assumption of normal distribution. Here, we can see that the residuals do not conform well to normality and randomness. Analyzing the reason, it may be due to the existence of individual outliers, which affects the distribution of the data and the distribution of the residuals, making them not conform to normality and randomness.**

**if github md knit fail,change to the html output“**

# Results

## Gender Differences in HIV Prevalence

A Welch Two Sample t-test was conducted to compare the prevalence of HIV diagnoses between female and male populations. The analysis revealed a statistically significant difference in the number of HIV diagnoses between females (\(\bar{x} = 6.14\)) and males (\(\bar{x} = 26.94\)), \(t(1921.9) = -8.183\), \(p < .00001\). The 95% confidence interval for the difference in means ranged from -25.78 to -15.81, indicating a higher prevalence of HIV diagnoses in males compared to females.

## HIV Diagnoses Across Ethnicities

An Analysis of Variance (ANOVA) was performed to examine whether the number of HIV diagnoses significantly differed across various ethnic groups. The results indicated a significant effect of race on the number of HIV diagnoses, \(F(5, 5173) = 59.623\), \(p < .00001\). This finding suggests substantial disparities in HIV diagnoses among different racial groups.

## HIV Prevalence Among Different Age Groups

Further, an ANOVA was conducted to assess the impact of age on HIV diagnoses. The analysis showed a significant difference in the number of HIV diagnoses among different age groups, \(F(6, 10569) = 178.22\), \(p < .00001\). This result highlights the varying prevalence of HIV across age demographics.


These findings collectively offer valuable insights into the epidemiological patterns of HIV diagnoses in New York City, highlighting significant disparities based on gender, race, and age. Such disparities underscore the need for targeted public health interventions and policies to address the unique needs of these demographic groups.

# Discussion

## Findings from the Analysis

 * Gender and HIV Prevalence: There tends to be a difference in HIV infection rates and diagnosis numbers between males and females.In the analysis we conducted, it was found that the number of HIV diagnoses was significantly higher in males compared to females. 
 * Ethnic Disparities in HIV Diagnoses: Our ANOVA tests revealed significant differences in HIV diagnoses among different races and age groups,emphasizing the persistence of inequities in HIV prevalence.
 * Variation Across Boroughs: The linear regression analysis showed significant differences in HIV diagnoses among the boroughs of New York City. Boroughs like Staten Island exhibited the largest negative coefficients in the model, indicating a lower number of HIV diagnoses compared to other regions. The differences are possibly influenced by factors such as local public health policies, access to healthcare services, and community-level prevention efforts.
 * Trends in HIV Infections: Our analysis indicated an 31% decrease in new HIV infections from 2017 to 2021, suggesting a gradual but steady progress in controlling new infections.
 
## Insights and Reflections
Our findings provide critical insights into the multifaceted nature of HIV epidemiology in New York City. The results underscore the necessity of addressing social and structural factors, such as racism, sexism, economic factors, homophobia, and transphobia. These determinants can influence risk behaviors, access to healthcare, and overall health outcomes.

While some findings, such as the disparities in HIV diagnoses among different demographics, were expected, the impact of external factors like the COVID-19 pandemic on the patterns of HIV diagnoses was particularly intriguing. It highlights the dynamic nature of public health challenges and the need for adaptive and resilient healthcare systems.


The insights gained from this analysis can aid in the development of more nuanced and effective public health strategies. These strategies should not only be responsive to the unique needs of different demographics but also adaptable to the specific challenges faced by each borough and race. By addressing these multifaceted aspects, New York City can continue to make progress towards reducing HIV infections and improving outcomes for people living with HIV.

