---
title: "Statistical Models"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(plotly)
library(ggplot2)

#input the New York HIV data from NYC opendata 
nyc_HIV1=
  GET("https://data.cityofnewyork.us/resource/ykvb-493p.csv",
  query = list("$limit" = 10000)) |>
  content("parsed")|>
  filter(year >= 2017)


nyc_HIV2=
  GET("https://data.cityofnewyork.us/resource/dxnu-p2qd.csv",
  query = list("$limit" = 22000)) |>
  content("parsed")|>
  filter(year >= 2017)

HIV_service=
  GET("https://data.cityofnewyork.us/resource/pwts-g83w.csv",
  query = list("$limit" = 1000)) |>
  content("parsed")|>
  select(unique_id,facilityname,service_type, address,borough,state,zipcode,latitude,longitude)
  

colnames(nyc_HIV1)
colnames(nyc_HIV2)
head(nyc_HIV1)
```

# Exploratory statistical analyses.
**Correlation analysis:After the Visualizations is there any association between different variables such as neighborhood, age, gender, and race.**

**T-test for HPV prevalence in gender-specific populations.**
```{r}
nyc_HIV1$hiv_diagnoses_num <- as.numeric(nyc_HIV1$hiv_diagnoses_num)

# Extracting data for females and males
female_data <- na.omit(nyc_HIV1[nyc_HIV1$sex == "Female",]$hiv_diagnoses_num)
male_data <- na.omit(nyc_HIV1[nyc_HIV1$sex == "Male",]$hiv_diagnoses_num)

# Perform T-test
t.test(female_data, male_data)
```
summary:
The t-test passed the test of significance indicating that there was a significant difference in the number of males and females with HIV.

**Test whether "hiv_diagnoses_num" is significantly different across ethnicities.**
```{r}
anova1 <- anova(lm(hiv_diagnoses_num ~ race, data = nyc_HIV1))
print(anova1)
```
summary:
The ANOVA passed the test of significance, indicating that there is indeed a significant difference in the number of people of different races who have HIV.

**Test whether "hiv_diagnoses_num" is significantly different between age groups.**
```{r}
anova2 <- anova(lm(hiv_diagnoses_num ~ age, data = nyc_HIV2))
print(anova2)
```
summary:
The ANOVA passed the test of significance indicating that there is a significant difference in the number of HIV in different age groups.

**Test whether there is a significant difference between people of different races and different age groups.**
```{r}
race_table <- table(nyc_HIV1$race, nyc_HIV1$sex)
race_table
# Perform chi-square test
chisq.test(race_table)
```
summary:
The chi-square test passes the test of significance, indicating that there is a significant relationship between race and gender.

# Linear regression analysis
```{r}
# Defining a Unique Thermal Coding Processor
library(caret)
encoder <- dummyVars(~ borough + sex + race, data = nyc_HIV1)
# Solo thermal coding of data
encoded_data <- predict(encoder, newdata = nyc_HIV1)
# Merging the coded and original data sets
nyc_HIV3 <- cbind(nyc_HIV1, encoded_data)
nyc_HIV4<-nyc_HIV3[,-c(2,3,4,5,7,8,9,10,11,12,18,21)]
model <- lm(hiv_diagnoses_num ~. , data = nyc_HIV4)
summary(model)
```
summary:
The coefficients of all variables passed the significance test, indicating that all of these independent variables have a significant effect on hiv_diagnoses_num. The coefficient before the variable "boroughStaten\\\nIsIand" has the largest absolute value, which means that it has the largest effect on the number of HIV patients. In addition, the coefficients before these variables are all negative, indicating that they all negatively affect hiv_diagnoses_num.Finally, the multiple linear regression model constructed ishiv_diagnoses_num=4514.4981-2.0224*year-354.5227*boroughBronx-358.4969*boroughBrooklyn-362.8923*boroughManhattan-363.2155*boroughQueens-373.2117*boroughStaten\\nIsland-26.8822*sexFemale-6.1453*sexMale-62.6910*raceAsian /Pacific\\nIslander-36.4375*raceBlack-41.7176*raceLatino/Hispanic-65.3393*raceOther/Unknown-57.9514*raceWhite.

# Regression model test
```{r}
#Drawing demonstration
par(mfrow=c(2, 2))  
plot(model)

# Plot of predicted values against residuals
residuals <- resid(model)
fitted_values <- predict(model)
plot(fitted_values, residuals,
     xlab = "Fitted values",
     ylab = "Residuals",
     main = "Residuals vs. Fitted")
abline(h = 0, col = "red")
```

**summary:Here is a test of the effectiveness of the regression model constructed above. Use plot(model) to draw residual plots, scatter plots and other diagnostic plots. Residual plots are used to check for linear properties, heteroskedasticity, outliers, etc. Scatter plots can be used to check for correlation between independent variables, and QQ plots are used to check whether the residuals conform to the assumption of normal distribution. Here, we can see that the residuals do not conform well to normality and randomness. Analyzing the reason, it may be due to the existence of individual outliers, which affects the distribution of the data and the distribution of the residuals, making them not conform to normality and randomness.**

**if github md knit fail,change to the html output“**
