---
title: "Data Processing and Exploratory analyses"
author: 
  - "Jake W. Coldiron,"
  - "Zixuan Qiu,"
  - "Tianyuan Deng,"
  - "Aung Nay Win,"
  - "Yuqi Cheng"
output: github_document
---

**Date:** 2023-11-10

```{r message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(plotly)
```

# Data processing

```{r message=FALSE}
#input the New York HIV data from NYC opendata 
nyc_HIV1=
  GET("https://data.cityofnewyork.us/resource/ykvb-493p.csv",
  query = list("$limit" = 10000)) |>
  content("parsed")|>
  filter(year >= 2017)


nyc_HIV2=
  GET("https://data.cityofnewyork.us/resource/dxnu-p2qd.csv",
  query = list("$limit" = 22000)) |>
  content("parsed")|>
  filter(year >= 2017)
  
```

```{r message=FALSE}
#HIV service resource 
HIV_service=
  GET("https://data.cityofnewyork.us/resource/pwts-g83w.csv",
  query = list("$limit" = 1000)) |>
  content("parsed")|>
  select(unique_id,facilityname,service_type, address,borough,state,zipcode,latitude,longitude)
  
```
**A short  clear description of where and how data were obtained. explain the variables means in the data**
**Calculate the sum, mean, median, standard deviation of HIV/AIDS indicators** 


# Visualizations

**Shows the number of HIV/AIDS diagnoses by race, sex, and age group.**
```{r}
#work
#work
colnames(nyc_HIV1)
colnames(nyc_HIV2)
head(nyc_HIV1)
library(ggplot2)
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

Female <- nyc_HIV1 %>% 
  filter(sex == "Female" & borough == "All" & race !="All")

Male <- nyc_HIV1 %>% 
  filter(sex == "Male"  & borough == "All" & race !="All")

graph1 = bind_rows(Female, Male)

nyc_HIV1$hiv_diagnoses_num <- as.integer(nyc_HIV1$hiv_diagnoses_num)
nyc_HIV2$hiv_diagnoses_num <- as.integer(nyc_HIV2$hiv_diagnoses_num)

ggplot(graph1, aes(x=race, y=hiv_diagnoses_num, fill=sex)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Race", y = "Number of HIV/AIDS Diagnoses", fill = "Sex")

# Graph Age 
age_summary <- nyc_HIV2 %>%
  filter(age != "All" & race == "All" & borough == "All") %>%
  group_by(age) %>%
  summarise(total_hiv_diagnoses = sum(hiv_diagnoses_num))

ggplot(age_summary, aes(x = age, y = total_hiv_diagnoses)) +
  geom_bar(stat = "identity", color = "yellow", width = 0.7) +
  labs(title = "Total HIV Diagnoses by Age Group",
       x = "Age Group",
       y = "Total HIV Diagnoses") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
summary:

**Show the map of NYC HIV service resources location.**
```{r}
if (!require(leaflet)) install.packages('leaflet')
library(leaflet)
filtered_data = HIV_service[HIV_service$longitude != 0 & HIV_service$latitude != 0,]
# Create the map
m <- leaflet(filtered_data) %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    popup = ~paste("<strong>Facility Name:</strong>", facilityname,
                   "<br><strong>Service Type:</strong>", service_type,
                   "<br><strong>Address:</strong>", address,
                   "<br><strong>Borough:</strong>", borough,
                   "<br><strong>Zip:</strong>", zipcode)
  )

# Print the map
print(m)
```
summary:

**Show the distribution of the HIV/AIDS diagnoses by neighborhood.** 
```{r}
#work
# I'll see if I can find the log&lat for the neighborhoods
diagnoses <- nyc_HIV1 %>% 
  select(borough, hiv_diagnoses_num)

# Plot bar chart
ggplot(diagnoses, aes(x = reorder(borough, hiv_diagnoses_num), y = hiv_diagnoses_num)) +
  geom_col() +
  labs(x = "Neighborhood",
       y = "Number of HIV Diagnoses",
       title = "HIV Diagnoses by Neighborhood") +
  coord_flip()
```
summary:

**Trend analysis:analyze the trend of the number of HIV/AIDS diagnoses over time.**
```{r}
#work
#work
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

# Plot diagnoses over time
ggplot(All, aes(x = year, y = hiv_diagnoses_num)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "HIV Diagnoses Trend (All Sex)",
    x = "Year",
    y = "Number of Diagnoses"
  )
```
summary:

# Exploratory statistical analyses.
**Correlation analysis:After the Visualizations is there any association between different variables such as neighborhood, age, gender, and race.**

```{r}
#work
```
T-test /ANOVA: Compare differences in HIV between two or more groups ( gender or racial).
Chi-square test: Analyze associations between categorical variables (gender, race).
